---
title: "Homework3"
author: "Yilin"
date: "3/18/2020"
output:
  md_document:
    variant: markdown_github
editor_options:
  chunk_output_type: inline
---


```{r}
library('MASS') ## for 'mcycle'
library('manipulate')
library('splines') ## 'ns'
library('caret') ## 'knnreg' and 'createFolds'
```

1. Randomly split the mcycle data into training (75%) and validation (25%) subsets.

```{r}
smp_size <- floor(0.75 * nrow(mcycle))
set.seed(123)
train_ind <- sample(seq_len(nrow(mcycle)), size = smp_size)
training <- mcycle[train_ind, ]
validation <- mcycle[-train_ind, ]
```

2. Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values.

```{r}
y <- training$accel
x <- matrix(training$times, length(training$times), 1)
```

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```

```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern,...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}
```


```{r create predictions}
k <- seq(1,50,1) #set tuning parameter k from 1 to 50
matrix_pred <- matrix(ncol = 50, nrow = nrow(training))
for(i in seq_along(k))
{
  y_hat <- nadaraya_watson(y, x, x, kern=kernel_k_nearest_neighbors, k = k[i])
  matrix_pred[,i] <- y_hat
}

#matrix_pred
```



3. With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d

effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}

```

```{r}
training_error <- rep(NA,50)
aic_error <- rep(NA,50)
bic_error <- rep(NA,50)
df_k <- rep(NA,50)
for (i in 1:50){
  df_k[i] <- effective_df(y, x, kern = kernel_k_nearest_neighbors, k = k[i])
}

```

```{r train error}
for (i in 1:50){
  training_error[i] <- error(y, matrix_pred[,i])
  aic_error[i] <- aic(y, matrix_pred[,i],df_k[i])
  bic_error[i] <- bic(y, matrix_pred[,i],df_k[i])
}
```

```{r}
y_validation <- validation$accel
x_validation <- matrix(validation$times, length(validation$times), 1)
y_hat_validation <- matrix(ncol = 50, nrow = nrow(validation))
for (i in (1:50)){
  validation_pred <- nadaraya_watson(y, x, x_validation,
  kern=kernel_k_nearest_neighbors, k=k[i])
  y_hat_validation[,i] <- validation_pred
}

#y_hat_validation
```

```{r validation error}
validation_error <- rep(NA,50)
for (i in 1:50){
  validation_error[i] <- error(y_validation,y_hat_validation[,i])
}

```

```{r plot}
plot(k, training_error, type='l', col = 'black',ylab = 'error')
lines(k,aic_error,type = 'l', col = 'green')
lines(k,bic_error, type = 'l', col = 'blue')
lines(k, validation_error, type = 'l', col = 'red')
text(30,1400,'AIC, BIC, Traning Error', col = 'blue')
text(30,600, 'Validation Error', col = 'red')
```

AIC, BIC and traning error seem to be the same line but they are slightly different from each other.

4. For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}
## 5-fold cross-validation of knnreg model
## create five folds
set.seed(1985)
m_flds  <- createFolds(mcycle$accel, k=5)
sapply(m_flds, length)  ## not all the same length
```

```{r}
cvknnreg <- function(kNN, flds=m_flds) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    m_trn <- mcycle[-flds[[tst_idx]],]
    m_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=m_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, m_tst)
    cverr[tst_idx] <- mean((m_tst$accel - pre_tst)^2)
  }
  return(cverr)
}
```

```{r}
## Compute 5-fold CV for kNN = 1:50
cverrs <- sapply(1:50, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)
```

5. Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}
## Plot the results of 5-fold CV for kNN = 1:50
cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)
plot(x=1:50, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:50, x1=1:50,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)
```

6. Interpret the resulting figures and select a suitable value for the tuning parameter.

From the figure, the tuning parameter K=14 will generate smallest CV-estimated test error. To avoid a complex modle, after applying one standard error rule, K=3 will be a sutiable value for the tuning parameter.


