---
title: "Homework4"
author: "Yilin"
date: "03/31/2020"
output:
  md_document:
    variant: markdown_github
---

```{r}
library(randomForest)
library(ElemStatLearn)
library(caret)
```


## Convert the response variable in the “vowel.train” data frame to a factor variable prior to training, so that “randomForest” does classification rather than regression.

```{r}
vowel.train$y = factor(vowel.train$y)
```

## Review the documentation for the “randomForest” function. Fit the random forest model to the vowel data using all of the 11 features using the default values of the tuning parameters.

```{r}
rf <- randomForest(y~.,data=vowel.train)
```

## Use 5-fold CV and tune the model by performing a grid search for the following tuning parameters: 1) the number of variables randomly sampled as candidates at each split; consider values 3, 4, and 5, and 2) the minimum size of terminal nodes; consider a sequence (1, 5, 10, 20, 40, and 80).

```{r}
mtry = c(3,4,5)
nodesize = c(1,5,10,20,40,80)
params = expand.grid(mtry = mtry,nodesize = nodesize)
params
```

```{r}
folds_5 <- createFolds(vowel.train$y,list = F, k = 5)
folds_5
```

```{r}
## use matrix to store all CV error
cv_error <- matrix(0,18,5)
for (i in 1:18){
  for (j in 1:5){
    train_data <- vowel.train[which(folds_5!=j),]
    test_data <- vowel.train[which(folds_5==j),]
    rf_grid <- randomForest(y ~.,data = train_data,
                            mtry = params[i,1],
                            nodesize = params[i,2])
    rf_pred <- predict(rf_grid, test_data)
    cv_error[i,j] <- mean(rf_pred != test_data$y)
  }
}
```

```{r}
mean_cv_tune <- apply(cv_error,1,mean)
min_cv <- min(mean_cv_tune)
ind <- which(mean_cv_tune == min_cv)
params[ind,]
```

Thus mtry =3, nodesize =1 will generate smallest cross validation error

## With the tuned model, make predictions using the majority vote method, and compute the misclassification rate using the ‘vowel.test’ data.

```{r}
vowel.test$y = factor(vowel.test$y)
```

```{r}
rf_tune <- randomForest(y ~.,data=vowel.train, mtry = 3, nodesize = 1)
rf_tune_pred <- predict(rf_tune, vowel.test)
mis_rate <- mean(rf_tune_pred != vowel.test$y)
mis_rate
```

